Stack Safety is a Security Property

### 2021-04-16 tex/main.tex
Tried to elucide confidentiality with more exampels

Next we consider {\em stepwise stack confidentiality}. There are a number of preliminary
concepts we must discuss here -- confidentiality quantifies over {\em variants} of the
initial state of a call, and requires that the traces from them move in {\em lockstep}.

@@

\definition Machine state pairs \(\mach,\mach'\) and \(\nach,\nach'\) are said to
\emph{change together}, written \(\mach,\mach' \diamond \nach,\nach'\), if,
for all components \(\component\) such that
\(\mach ~ \component \not = \mach' ~ \component\) or
\(\nach ~ \component \not = \nach' ~ \component\),
\(\mach' ~ \component = \nach' ~ \component\).

\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep}
if their adjacent pairs of states change together. We will represent this relation with
\(\machT \doteq \nachT\). \bcp{People may not be familiar with the notation}
\sna{It's a new notation, we use this enough that it seems worth picking an
  evocative symbol.}\bcp{Sorry -- I meant the notational convention of representing a
coinductive relation with inference rules in this way, not the symbol...}
We define it as a coinductive relation:

\begin{minipage}{.3\textwidth}
\judgment{}
         {\(\mach \doteq \nach\)}
\end{minipage}
\begin{minipage}{.6\textwidth}
\judgmenttwo{\(\mach, \head(\machT) \diamond \nach, \head(\nachT)\)}
            {\(\machT \doteq \nachT\)}
            {\(\mach \machT \doteq \nach \nachT\)}
\end{minipage}

\definition For any depth \(\depth\), define \(f_\depth ~ \mach ~ (\_, \rts)\) to be
\(|\rts| \geq \depth\). Then each reachable segment \(\MPCT \langle f_\depth \rangle\)
corresponds to a trace of a function from its call to its return.
Given such a segment \(S\), let \(\mach,\_,(\domMap,\_)\)  = \(\head(S)\) and let
\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \outside\).
A system enjoys \emph{stepwise stack confidentiality} if, for all \(\MPCT\langle f_\depth\rangle\)
with associated \(\mach\) and \(\components\) as above, and
for all \(\nach\) such that \(\mach \approx_\components \nach\) and
 \(\nach \hookrightarrow \NPCT\),
 \(\MPCT\langle f_\depth\rangle \doteq \NPCT | \neg f_\depth \).\apt{I found the previous
   version hopelessly confusing, but I'm not at all sure that this one is right! SNA please check.}
 \sna{It's right, but I need to rethink how the M<f> notation works...}

@@

\section{Do Micro-Policies Enforce Stack Safety?}
\label{sec:enforcement}

\bcp{``Enforcement'' seems like too broad a title.  What about calling it
  ``Evaluating Micro-Policies'' or ``Do Micro-Policies Enforce Stack
  Safety?''? And what about folding the testing section into this one?  That
would keep related things together, would avoid appearing to make overly
grand claims in either section, and would give us an excuse to write less
about the mechanism here---we only need to write enough about the mechanism
itself that readers can understand our validation case study.}
\leo{Agreed. Took a crack at it.}

@@

In this section we sketch an existing enforcement mechanism, Depth
Isolation from \citet{DBLP:conf/sp/RoesslerD18}, that implements the
formal stack-safety property described in the previous section; in
fact, it enforces a stronger, inductive version of that property which
strengthens the local state encapsulation component by checking for
integrity and confidentiality violations at every step of the callee's
execution trace. We validate that claim by developing a property-based
testing framework that can quickly uncover inconsistencies between the
enforcement mechanism and the stack-safety property.

\subsection*{A Conservative Stack-Safety Enforcement Mechanism}
%
The enforcement policy of \citeauthor{DBLP:conf/sp/RoesslerD18} relies
on a programmable, tag-based reference monitor that runs alongside the
program.  To control this monitor, they use a programming model that
allows fine-grained manipulation of metadata tags to encode so-called
\emph{micro-policies}~\citep{pump_oakland2015}.
In such a system, all values in
memory addresses and registers (including the $\PCname$)
are enriched with an abstract metadata tag, which can represent
arbitrary information about the value. A micro-policy is defined as a set of tags and a
collection of software-defined rules, indexed by machine opcode.
At each step of the machine, the relevant rule is applied to the tags on
the instruction's inputs ($\PCname$, registers, memory) and on the instruction itself, and produces
one of two outcomes: either the instruction is allowed to execute (and generates
tags for the result of the operation and the new $\PCname$), or the machine
fail-stops with a policy violation.
\citeauthor{pump_oakland2015}\bcp{and PIPE, if there are any references to
  cite} have shown that a wide range of micro-policies can
be defined using this scheme.

@@

% \leo{I'm not a fan of the next paragraph.}\bcp{Yes, it's pretty far down in
%   the weeds.  I don't think this level of detail is needed, in the context
%   of a paper that is mostly about our definitions, not theirs.}
% Note that in this policy both the entry and exit sequences
% must write to each element of
% the frame, which can be quite expensive, especially for programs that
% allocate large but sparsely populated frames; we return to this point
% in \cref{sec:lazy}.
% %
% For example, the program in \cref{fig:passingsimple} manages the stack
% in this way. Here, the call pseudo-instruction to {\tt f} in line 5
% hides a blessed sequence of instructions that allocate space to save
% its return address to {\tt main} and seal the protected part of the
% stack before {\tt f}'s body starts executing at 10. Conversely, the
% return instruction in line 15 restores and jumps to the return address
% and deallocates its frame space, and unseals the corresponding parts
% of the stack. The function itself allocates and deallocates its own
% private frame space in instructions 10 and 14. Additionally, unlike in
% the simplified presentation of the policy, the caller shares the
% arguments passed to the callee through the stack, and for this creates
% and initializes (3--4) and destructs (6) the argument part of the
% callee's stack frame.

%\rb{Note that we have also considered and implemented more flexible schemes
%  based on allowing operations above the current activation depth. How do we put
%  these together?}
%\rb{Do we want to discuss variations involving frame pointers or more implicit
%  return addresses, argument passing on the stack, alloc/dealloc, etc.?}
%\rb{Where to talk about the role of the compiler?}

@@

\subsection*{Validation through Random Testing}
\label{sec:testing}

@@

Naturally, testing also revealed a number of errors in our
implementation of the enforcement mechanism (the original was written in C++
and targeted ARM machine code
%\bcp{right?}\leo{yeah}
; we re-implemented it in Coq targeting
RISC-V).  These errors ranged from trivial typos to
ones that required an
intriguingly complex setup to appear.  Notably, our first try at the
``blessed'' call sequence/policy combination did not initialize all
locations for the newly allocated stack frame correctly, but left some
of them as $\tagNoDepth$. This allowed for a potential integrity
violation, but only if a rather complicated sequence of events
occured: \leo{call - write - call (seal) - return - call - write?}

>>>>

Next we consider {\em stepwise stack confidentiality}. Recall that our properties are
nested: we will quantify over subtraces that represent whole calls. For each such
subtrace, we will quantify over {\em variants} of its initial state. Then we require
that the original, {\em primary} traces move in {\em lockstep} with the variant.

\definition \(\MPCT^*\) is the set of all traces of calls, defined as
\[\begin{align*}
\bigcup_{\depth \in \mathbb{Z}} \{\MPCT | \MPCT\langle f_\depth \rangle
\text{ is a reachable segment}\} \hspace{1em} &
\text{where } f_\depth ~ (\mach,\pol,(\domMap,\rts)) \triangleq |\rts| \geq \depth
\end{align*}\]

@@

\definition Let \(\Delta(\mach,\mach')\) be the set of set of components \(\component\)
such that \(\mach ~ \component \not = \mach' ~ \component\).

\definition Between two pairs of states \((\mach,\mach')\) and \((\nach,\nach')\), we
say that they {\em lock} if for all components \(\components\) in
\(\Delta(\mach,\mach') \cup \Delta(\nach,\nach')\),
\(\mach' ~ \component = \nach' ~ \component\). We write this \((\mach,\mach') \diamond
(\nach,\nach')\).

\definition A pair of traces \(\machT\) and \(\nachT\) are {\em in lockstep}
if their adjacent pairs of states change together. We will represent this relation with
\(\machT \doteq \nachT\). We define it as a coinductive relation:

\begin{minipage}{.3\textwidth}
\judgment{}
         {\(\mpcstatename \doteq \npcstatename\)}
\end{minipage}
\begin{minipage}{.6\textwidth}
\judgmenttwo{\(\mach, \pi_\mach(\head(\machT)) \diamond \nach, \pi_\mach(\head(\nachT))\)}
            {\(\machT \doteq \nachT\)}
            {\(\notfinished{(\mach,\_,\_)}{\machT} \doteq \notfinished{(\nach,\_,\_)}{\nachT}\)}
\end{minipage}

\definition {\bf Stepwise Stack Confidentiality}

For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
let \(\components\) be the set of components \(\component\) such that
\(\domMap ~ \component = \outside\).
A system enjoys \emph{stepwise stack confidentiality} if,
for all \(\nach\) such that \(\mach \approx_\components \nach\) and
\(\nach \hookrightarrow \NPCT | \neg f_{|\rts|}\),
\(\MPCT \doteq \NPCT\).

\vspace{\afterdisplayskip}

Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
of the components in \(\components\), and in addition their subtraces are also in
\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
will contain these calls and returns as well, so its return stack pushes and pops
in lockstep until it breaks at the final return.

@@

\section{Enforcement}
\label{sec:enforcement}

\bcp{``Enforcement'' seems like too broad a title.  What about calling it
  ``Evaluating Micro-Policies'' or ``Do Micro-Policies Enforce Stack
  Safety?''? And what about folding the testing section into this one?  That
would keep related things together, would avoid appearing to make overly
grand claims in either section, and would give us an excuse to write less
about the mechanism here---we only need to write enough about the mechanism
itself that readers can understand our validation case study.}

@@

In this section we sketch an existing enforcement mechanism,
Depth Isolation from \citet{DBLP:conf/sp/RoesslerD18}, that implements
the formal stack-safety property described in the previous section; in
fact, it enforces a stronger, inductive version of that property which strengthens the
local state encapsulation component by checking for integrity and
confidentiality violations at every step of the callee's execution trace. \bcp{Examples?}

%\paragraph*{Micro-Policies}
%
The enforcement policy of \citeauthor{DBLP:conf/sp/RoesslerD18} relies
on a programmable, tag-based reference monitor that runs alongside the
program.  To control this monitor, they use a programming model that
allows fine-grained manipulation of metadata tags to encode so-called
\emph{micro-policies}~\citep{pump_oakland2015}.
In such a system, all values in
memory addresses and registers (including the $\PCname$)
are enriched with an abstract metadata tag, which can represent
arbitrary information about the value. A micro-policy is defined as a set of tags and a
collection of software-defined rules, indexed by machine opcode.
At each step of the machine, the relevant rule is applied to the tags on
the instruction's inputs ($\PCname$, registers, memory) and on the instruction itself, and produces
one of two outcomes: either the instruction is allowed to execute (and generates
tags for the result of the operation and the new $\PCname$), or the machine
fail-stops with a policy violation.
\citeauthor{pump_oakland2015}\bcp{and PIPE, if there are any references to
  cite} have shown that a wide range of micro-policies can
be defined using this scheme.

@@

\leo{I'm not a fan of the next paragraph.}\bcp{Yes, it's pretty far down in
  the weeds.  I don't think this level of detail is needed, in the context
  of a paper that is mostly about our definitions, not theirs.}
Note that in this policy both the entry and exit sequences
must write to each element of
the frame, which can be quite expensive, especially for programs that
allocate large but sparsely populated frames; we return to this point
in \cref{sec:lazy}.
%
For example, the program in \cref{fig:passingsimple} manages the stack
in this way. Here, the call pseudo-instruction to {\tt f} in line 5
hides a blessed sequence of instructions that allocate space to save
its return address to {\tt main} and seal the protected part of the
stack before {\tt f}'s body starts executing at 10. Conversely, the
return instruction in line 15 restores and jumps to the return address
and deallocates its frame space, and unseals the corresponding parts
of the stack. The function itself allocates and deallocates its own
private frame space in instructions 10 and 14. Additionally, unlike in
the simplified presentation of the policy, the caller shares the
arguments passed to the callee through the stack, and for this creates
and initializes (3--4) and destructs (6) the argument part of the
callee's stack frame.
%
%\rb{Note that we have also considered and implemented more flexible schemes
%  based on allowing operations above the current activation depth. How do we put
%  these together?}
%\rb{Do we want to discuss variations involving frame pointers or more implicit
%  return addresses, argument passing on the stack, alloc/dealloc, etc.?}
%\rb{Where to talk about the role of the compiler?}

@@

\section{Randomized Testing}
\label{sec:testing}

@@

Naturally, testing also revealed a number of errors in our
implementation of the enforcement mechanism (the original was written in C++
and targeted ARM machine code\bcp{right?}; we re-implemented it in Coq targeting
RISC-V).  These errors ranged from trivial typos to
ones that required an
intriguingly complex setup to appear.  Notably, our first try at the
``blessed'' call sequence/policy combination did not initialize all
locations for the newly allocated stack frame correctly, but left some
of them as $\tagNoDepth$. This allowed for a potential integrity
violation, but only if a rather complicated sequence of events
occured: \leo{call - write - call (seal) - return - call - write?}
